% ************************** Thesis Abstract *****************************
% Use `abstract' as an option in the document class to print only the titlepage and the abstract.
\begin{abstract}
Metabolites and reactions taking part in Lipid metabolism processes
are often poorly annotated. This makes it difficult to analyse with
current traditional techniques used for metabolic processes like Flux
Balance Analysis (FBA). Moreover because of the low number of carbons flowing
towards Lipid metabolism pathways the effect of the inherent
probabilistic nature of chemical events is amplified. This behaviour
is then not very well described by the average deterministic view
offered by FBA. Here we propose an alternative stochastic
reaction-centric view of Lipid metabolism to
better capture some of its defining characteristics on a more local
scale: iterative conversion processes, probabilistic decisions, and
crucial regulatory mechanisms. Regulatory mechanisms are thought to be
tightly regulated and their disruption leads to loss of metabolic
balance that seems to be consistent with disorders. 
We also assess the formal graphical
language of Petri Nets to capture this stochastic reaction-centric
view and contrast it with the pi-calculus process algebra that is
another popular approach for executable models in Biology.  Here we
particularly focus on Fatty Acid (FA) synthesis that is a core part of
lipid metabolism. A basic high-level Petri Net and stochastic
pi-calculus model for the FA
synthesis/elongation presented along with techniques for tuning it with
experimental datasets. The Petri Net model is then further extended to include
an important  regulatory mechanism of FA synthesis activity through
the action of AMPK protein kinase.
\end{abstract}
